Synthesising Termination Proofs for Bit-Vector Programs

### 2014-07-06 papers/termination/paper.tex
nudge

To evaluate our algorithm, we implemented two tools: one to generate a termination specification
from a C program, and one to synthesise a proof from the specification.  We ran the resulting
termination prover on 47 benchmarks taken from the literature and SV-COMP'15~\cite{svcomp}.  To provide a comparison
point, we also ran {\sc ARMC}~\cite{armc-website} on the same benchmarks.  Each tool was given a time limit of
180s, and was run on an unloaded 8-core 3.07\,GHz Xeon X5667 with 50\,GB of RAM.  The
results of these experiments are shown in Figure~\ref{fig:experiments}.

It should be noted that the comparison here is imperfect, since {\sc ARMC} is solving a slightly different
problem -- it checks whether the program under analysis would terminate if run with unbounded integer variables,
while we are checking whether the program terminates with fixed width integers variables.  This means that
{\sc ARMC}'s verdict differs from ours in 3 cases (due to the differences between integer and bit-vector semantics).
There are a further 5 cases where our tool is able to find a proof and {\sc ARMC} cannot, which we believe is due
to our more expressive proof language.  In 3 cases, {\sc ARMC} times out while our tool is able to find a termination
proof.  All of these cases have nested loops, which are easy for us to handle, but can be difficult for
provers based on enumerating lassos.

@@

The final column in Figure~\ref{fig:experiments} shows the number of iterations of the CEGIS loop our
synthesiser performed.  While the theoretical bound for the number of CEGIS iterations is very large
(exponential in the amount of space used by the program), these results demonstrate that in practice
the number of iterations required is usually small.

@@

 \section{Conclusions and Related Work}
%\subsection{Termination analysis}
Automated program termination is a research topic that has received a fair amount of attention from the software verification community.
In order to compare our technique to the rest of the area, 
Figure~\ref{fig:handletable} summarises the related works with respect to the assumptions they make about programs and ranking functions.

>>>>

To evaluate our algorithm, we implemented two tools: one to generate a
termination specification from a C program, and one to synthesise a proof
from the specification.  We ran the resulting termination prover on 47
benchmarks taken from the literature and SV-COMP'15~\cite{svcomp}.  To
provide a comparison point, we also ran {\sc ARMC}~\cite{armc-website} on
the same benchmarks.  Each tool was given a time limit of $180\,s$, and was
run on an unloaded 8-core 3.07\,GHz Xeon X5667 with 50\,GB of RAM.  The
results of these experiments are shown in Figure~\ref{fig:experiments}.

It should be noted that the comparison here is imperfect, since {\sc ARMC}
is solving a slightly different problem -- it checks whether the program
under analysis would terminate if run with unbounded integer variables,
while we are checking whether the program terminates with fixed-width
integer variables.  This means that {\sc ARMC}'s verdict differs from ours
in 3 cases (due to the differences between integer and bit-vector
semantics).  There are a further 5 cases where our tool is able to find a
proof and {\sc ARMC} cannot, which we believe is due to our more expressive
proof language.  In 3 cases, {\sc ARMC} times out while our tool is able to
find a termination proof.  All of these cases have nested loops, which are
easy for us to handle, but can be difficult for provers that enumerate
lassos.

@@

The final column in Figure~\ref{fig:experiments} gives the number of
iterations of the CEGIS loop our synthesiser performed.  While the
theoretical bound for the number of CEGIS iterations is very large
(exponential in the amount of space used by the program), these results
demonstrate that in practice the number of iterations required is usually
small.

@@

\section{Conclusions and Related Work}
%\subsection{Termination analysis}

Automated program termination is a research topic that has received a fair
amount of attention from the software verification community.  In order to
compare our technique to the rest of the area, Figure~\ref{fig:handletable}
summarises the related works with respect to the assumptions they make about
programs and ranking functions.
